Quantifying Limits to Detection of Early Warning for Critical Transitions

### 2012-02-09 boettiger.tex
minor edit

 The ROC curves for these data (Fig.~\ref{fig:summary_roc}) show that the summary-statistic based indicators 
frequently lack the sensitivity to distinguish reliably between observed patterns from a stable or unstable system. 
The large correlations observed in the empirical examples (Fig.~\ref{fig:summary_stats}) are not uncommon in stable systems. 
For instance, we can read from the curve what fraction of true positives would be detected by a variance-based indicator
set to a 5\% false positive rate:
25\% in the simulated set,
5.0\% in the \emph{Daphnia} set,
and 5.4\% in the Glaciation data set.
Compare this to likelihood: 61\%, 34\%, and 100\% of the replicate true positives
would be correctly identified from the respective systems.
The value of the ROC curve is that we need not restrict the analysis to a pre-specified error such as 5\%,
as we would in classical hypothesis testing.  
We could instead decide that we wanted not to catch no less than 90\% of the true-positives,
and read off the corresponding false positives.  
For variance this corresponds respectively to a 49\%, 81\%, and 93\% false positive rate,
while for likelihood a 55\%, 35\%, 0\% false positive.

>>>>

 The ROC curves for these data (Fig.~\ref{fig:summary_roc}) show that the summary-statistic based indicators 
frequently lack the sensitivity to distinguish reliably between observed patterns from a stable or unstable system. 
The large correlations observed in the empirical examples (Fig.~\ref{fig:summary_stats}) are not uncommon in stable systems. 
For instance, we can read from the curve what fraction of true positives would be detected by a variance-based indicator
set to a 5\% false positive rate:
25\% in the simulated set,
5.0\% in the \emph{Daphnia} set,
and 5.4\% in the Glaciation data set.
Compare this to likelihood: 61\%, 34\%, and 100\% of the replicate true positives
would be correctly identified from the respective systems.
The value of the ROC curve is that we need not restrict the analysis to a pre-specified error such as 5\%,
as we would in classical hypothesis testing.  
We could instead decide that we wanted no less than 90\% of the true-positives,
and read off the corresponding false positives.  
For variance this corresponds respectively to a 49\%, 81\%, and 93\% false positive rate, respectively,
while for likelihood this gives us respectivel 55\%, 35\%, and 0\% false positives.
